---
title: "EDA Titanic Dataset"
author: "Arshdeep Singh"
date: "06/10/2020"
output: html_document
---

## Introduction to the dataset

This dataset includes information about the passengers of RMS Titanic and also have info about is that particular passenger survived or not. For every individual person we have information about -

* **survived** - Person survived or not (0 for No and 1 for Yes) (Integer)
* **pclass** - With what class of ticket the passenger was traveling (in numbers) (Integer)
* **sex** - Male or Female (Character)
* **age** - Age of passenger in years (Double)
* **sibsp** - Number of siblings/spouse on Titanic (Integer)
* **parch** - Number of parent or children (Integer)
* **fare** - Amount of money person paid to travel (Double)
* **embarked** - Port of Embarkation (C = Cherbourg, Q = Queenstown, S = Southampton) (Character)
* **class** - With what class of ticket the passenger was travelling (in words) (Character)
* **who** - Man or Women or Child (Character)
* **adult_male** - Is the passenger an adult male (Logical)
* **deck** - Deck letter (9 levels A-G) (Character)
* **embark_town** - Port of Embarkation (Cherbourg, Queenstown, Southhampton) (Character)
* **alive** - Is the person alive (Character)
* **alone** - Was the person traveling alone (Logical) 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

## Install and load packages

``` {r}
library(data.table)
library(ggplot2)
```

## Import the dataset

```{r, eval = FALSE}
dataset <- fread("titanic.csv")
head(dataset)
```
```{r, echo = FALSE}
dataset <- fread("titanic.csv")
kable(head(dataset))
```

## Add NA to 'deck' feature
The 'deck' feature contains empty spaces instead of NA, so we substitute empty spaces with NA

```{r}
dataset$deck[dataset$deck == ""] <- NA
```

## Check Number of NA in each feature

```{r, comment = " "}
colSums(is.na(dataset))
```


## Summary of the dataset

```{r, comment = " "}
summary(dataset)
```


## Convert categorical features to factors

```{r}
dataset$pclass <- as.factor(dataset$pclass)
dataset$survived <- as.factor(dataset$survived)
dataset$sex <- as.factor(dataset$sex)
dataset$who <- as.factor(dataset$who)
dataset$embarked <- as.factor(dataset$embarked)
dataset$sibsp <- as.factor(dataset$sibsp)
dataset$parch <- as.factor(dataset$parch)
dataset$adult_male <- as.factor(dataset$adult_male)
dataset$embark_town <- as.factor(dataset$embark_town)
dataset$alive <- as.factor(dataset$alive)
dataset$alone <- as.factor(dataset$alone)
dataset$class <- as.factor(dataset$class)
dataset$deck <- as.factor(dataset$deck)
```

## Check structure of dataset

```{r, comment = " "}
str(dataset)
```

## Impute 'age' feature with respect to 'pclass'

```{r, warning = FALSE}
box_plot <- ggplot(dataset, aes(x = pclass, y = age), na.exclude = TRUE) + 
geom_boxplot() + 
scale_y_continuous(breaks = round(seq(min(dataset$age, na.rm = TRUE), 
                                      max(dataset$age, na.rm = TRUE), 
                                      by = 4), 
                                  1))

box_plot
```
```{r}
for(i in 1:length(dataset$age)){
    if(is.na(dataset$age[i])){
        if(dataset$pclass[i] == 1){
            dataset$age[i] <- 37
        }
        else if(dataset$pclass[i] == 2){
            dataset$age[i] <- 29
        }
        else{
            dataset$age[i] <- 24
        }
    }
}
```

## Survival rate with respect to sex

```{r}
ggplot(data = dataset, aes(x = sex, fill = survived)) + 
geom_bar() + 
scale_fill_discrete(name = "Survived", labels = c("No", "Yes")) +
scale_y_continuous(breaks = round(seq(0, 600, by = 50), 1))
```
```{r}
females_count = 0
males_count = 0
females_survived = 0
males_survived = 0

for(i in 1:length(dataset$sex)){
    if(dataset$sex[i] == 'male'){
        males_count = males_count + 1
        if(dataset$survived[i] == 1){
            males_survived = males_survived + 1
        }
    }
    else{
        females_count = females_count + 1
        if(dataset$survived[i] == 1){
            females_survived = females_survived + 1
        }
    }
}

print(paste("Total number of males: ", males_count))
print(paste("Total number of females: ", females_count))
print(paste("Number of males who survived: ", males_survived))
print(paste("Number of females who survived: ", females_survived))
print(paste("Percentage of males who survived: ", round(males_survived / males_count, 2)))
print(paste("Precentage of females who survived: ", round(females_survived / females_count, 2)))
```

## Survival rate with respect to class

```{r}
ggplot(data = dataset, aes(x = pclass, fill = survived)) +
geom_bar() + 
scale_fill_discrete(name = "Survived", labels = c("No", "Yes")) +
scale_y_continuous(breaks = round(seq(0, 500, by = 50), 1))
```
```{r}
class1_number = 0
class2_number = 0
class3_number = 0
class1_survived = 0
class2_survived = 0
class3_survived = 0

for(i in 1:length(dataset$pclass)){
    if(dataset$pclass[i] == 1){
        class1_number = class1_number + 1
        if(dataset$survived[i] == 1){
            class1_survived = class1_survived + 1
        }
    }
    else if(dataset$pclass[i] == 2){
        class2_number = class2_number + 1
        if(dataset$survived[i] == 1){
            class2_survived = class2_survived + 1
        }
    }
    else{
        class3_number = class3_number + 1
        if(dataset$survived[i] == 1){
            class3_survived = class3_survived + 1
        }
    }
}

print(paste("Total number of people in class 1: ", class1_number))
print(paste("Total number of people in class 2: ", class2_number))
print(paste("Total number of people in class 3: ", class3_number))
print(paste("Number of people in class 1 who survived: ", class1_survived))
print(paste("Number of people in class 2 who survived: ", class2_survived))
print(paste("Number of people in class 3 who survived: ", class3_survived))
print(paste("Percentage of people in class 1 who survived: ", round(class1_survived / class1_number, 2)))
print(paste("Percentage of people in class 2 who survived: ", round(class2_survived / class2_number, 2)))
print(paste("Percentage of people in class 3 who survived: ", round(class3_survived / class3_number, 2)))
```

## Plot 'age' with respect to 'survived' and 'who'

```{r}
ggplot(data = dataset, aes(x = age, y = survived,color = who)) +
geom_point()
```

## Plot 'age' with respect to 'survived' and 'sex'

```{r}
ggplot(data = dataset, aes(x = age, y = survived,color = sex)) +
geom_point()
```

## Plot 'age' with respect to 'survived' and 'pclass'

```{r}
ggplot(data = dataset, aes(x = age, y = survived,color = pclass)) +
geom_point()
```

## Survival rate with respect 'who'

```{r}
ggplot(data = dataset, aes(x = who, fill = survived)) +
geom_bar() + 
scale_y_continuous(breaks = round(seq(0, 550, by = 50), 1))
```
```{r}
men_count = 0
women_count = 0
children_count = 0
men_survived = 0
women_survived = 0
children_survived = 0

for(i in 1:length(dataset$who)){
    if(dataset$who[i] == "child"){
        children_count = children_count + 1
        if(dataset$survived[i] == 1){
            children_survived = children_survived + 1
        }
    }
    else if(dataset$who[i] == "man"){
        men_count = men_count + 1
        if(dataset$survived[i] == 1){
            men_survived = men_survived + 1
        }
    }
    else{
        women_count = women_count + 1
        if(dataset$survived[i] == 1){
            women_survived = women_survived + 1
        }
    }
}

print(paste("Total number of children: ", children_count))
print(paste("Total number of men: ", men_count))
print(paste("Total number of women: ", women_count))
print(paste("Number children who survived: ", children_survived))
print(paste("Number of men who  survived: ", men_survived))
print(paste("Number of women who  survived: ", women_survived))
print(paste("Percentage of children who survived: ", round(children_survived / children_count, 2)))
print(paste("Percentage of men who  survived: ", round(men_survived / men_count, 2)))
print(paste("Percentage of women who survived: ", round(women_survived / women_count, 2)))
```
















































































































